\section{Introduction}

Coding agents are now capable enough to generate substantial code, run tools, and execute multi-step workflows. The harder problem is no longer only code generation; it is oversight calibration. Developers are often pushed into one of two poor interaction modes: approve nearly every action, or grant broad autonomy and hope the agent stays within acceptable bounds. The first mode increases friction and stalls progress. The second increases the risk of subtle bugs, security mistakes, and changes the developer does not fully understand.

A preliminary formative survey we conducted with 21 experienced coding-agent users exposed this tension clearly. Respondents reported frequent concerns about subtle bugs, outdated context, security mistakes, and misaligned permission-seeking behavior. At the same time, they did not ask for constant oversight. The dominant preference was milestone-based or risk-based check-ins: pause for architectural tradeoffs, interface changes, or sensitive operations, but continue autonomously for routine work. Respondents also wanted visibility into why the system chose to interrupt or continue. These needs align with recent work on agent oversight and personalization, which highlights the difficulty of maintaining useful human control without turning oversight into a bottleneck~\\cite{grunde2026overseeing,huq2026cowcorpus,liang2026pahf}. They also align with the broader literature on appropriate reliance, which argues that human-AI systems should be designed to avoid both over-trust and under-trust rather than maximizing automation by default~\\cite{bansal2021appropriate}.

We argue that this is fundamentally a governance problem. Static rule files can encode some explicit preferences, but many preferences are implicit and only emerge through interaction: what a developer routinely corrects, how long they spend reviewing changes, which kinds of changes they wave through, and where they repeatedly insist on caution. Existing coding-agent systems do not provide a clean mechanism for converting those observations into inspectable future behavior while preserving a strong enforcement boundary around the local development environment.

This paper presents \emph{Smart Coder}, a governance-first framework for adaptive autonomy in coding agents. Smart Coder is built around a strict separation of responsibility. The model is untrusted and has no direct filesystem or shell authority. It can request reads, propose plans, surface check-ins, and generate candidate updates, but a local CLI retains authority over every read, write, verification step, and state transition. This lets autonomy be studied as an explicit policy layer over model behavior rather than an opaque property of the model itself.

The central mechanism in Smart Coder is a trace-driven adaptation loop. Each interaction produces structured traces that capture requested actions, affected files, user decisions, review timing, corrections, verification outcomes, and whether the interruption was initiated by the CLI or the model. These traces influence future behavior in two ways. First, they affect local policy decisions about whether future actions should proceed, be flagged, or require review. Second, they are summarized into dynamic system-prompt context that informs the model about high-trust areas, low-trust areas, relevance-ranked historical corrections, behavioral guidelines, autonomy mode, and workflow phase. The result is an externalized personalization loop that is inspectable, model-agnostic, and locally enforced.

Smart Coder also distinguishes between two initiators of oversight. The CLI may force a check-in when constraints, milestone triggers, or policy conditions indicate elevated risk. Separately, the model may proactively issue a check-in when it detects ambiguity, assumptions, or architectural tradeoffs. Because both are logged separately, Smart Coder supports a direct empirical question: which source of oversight is better calibrated under which conditions? This dual-initiator design is one of the system's primary research contributions.

The current prototype implements a heuristic baseline rather than a learned policy. This is deliberate. Our goal is not to claim an already-optimal autonomy policy, but to present a working artifact that demonstrates governed adaptive autonomy end to end and collects the traces needed for later calibration. In that sense, Smart Coder is both a usable interactive system and an experimental platform.

The contributions of this work are:
\begin{itemize}
    \item a governance-first architecture for coding agents in which the model is structurally untrusted and all side effects are mediated by a local CLI;
    \item a dual-initiator oversight design that supports both policy-initiated and model-initiated check-ins;
    \item a trace-driven adaptive autonomy loop that connects developer interaction history to both future policy decisions and future prompt construction; and
    \item a working prototype with milestone-aware check-ins, spec-aware planning, verification-aware autonomy adjustment, and exportable traces for future study.
\end{itemize}

This paper focuses on the system as built: its architecture, its current runtime behavior, and the demo scenario we use to illustrate its contribution. Formal user studies and learned calibration remain future work.
