\section{Demo Scenario and Evidence}

The demo is designed to show Smart Coder's contribution through a short, concrete two-session interaction on a small external codebase. Rather than demonstrating arbitrary code generation, the scenario emphasizes governed autonomy, model-side architectural reasoning, and cross-session adaptation.

\subsection{Demo task}

We use a small Python task API fixture with route handlers, service-layer validation, structured application errors, and a short task specification. The specification preserves response envelopes and handler signatures while allowing low-risk internal improvements. One file is marked as interface-sensitive, one path is explicitly locked, and the system is configured to run tests after edits.

The first session asks Smart Coder to add a summary view of task counts by status while preserving the existing API envelope and public handler signatures. This task is chosen because it creates a real architectural fork: either extend an existing handler or add a dedicated summary route. That fork is exactly the kind of decision Smart Coder is meant to surface through a model-initiated check-in.

The second session asks for a related follow-up change, such as priority filtering and tighter validation messaging, while preserving the same interface constraints. Between the two sessions, the developer's feedback from the first run remains in the trace database. The second run therefore demonstrates whether Smart Coder can change its behavior based on prior interaction rather than treating every session as a cold start.

\subsection{What attendees will see}

The live demonstration highlights four moments:
\begin{enumerate}
    \item \textbf{Spec-aware planning.} Smart Coder shows an intent declaration grounded in the provided task specification rather than immediately generating code.
    \item \textbf{A model-initiated architectural check-in.} The model surfaces a design tradeoff, states its assumptions, and presents explicit options instead of silently picking one.
    \item \textbf{Selective autonomy under governance.} Low-risk work proceeds with less friction, while interface-sensitive or constrained work still triggers review.
    \item \textbf{Cross-session adaptation.} A preference or correction given in the first session appears in the second session as prompt context, learned preference state, or altered oversight behavior.
\end{enumerate}

This sequence is the most compelling way to show Smart Coder's intended contribution: it is neither a static rules engine nor an unconstrained autonomous agent, but a governed system that adapts over time.

\subsection{Evidence in the current artifact}

The CAIS demo track asks authors to show evidence that the system works and is meaningfully different from alternatives. Smart Coder's current evidence is artifact-centric rather than a completed user study.

\begin{table}[t]
  \caption{Evidence to include in the final submission. Replace placeholder values with exported data from one or more representative pilot sessions.}
  \label{tab:evidence}
  \centering
  \begin{tabular}{p{0.44\columnwidth}p{0.42\columnwidth}}
    \toprule
    Evidence item & Current source \\
    \midrule
    Representative task success & \textbf{Placeholder:} exported session bundle \\
    Number of model-initiated check-ins & \textbf{Placeholder:} \texttt{observe checkin-stats} \\
    Number of policy-initiated check-ins & \textbf{Placeholder:} \texttt{observe checkin-stats} \\
    Cross-session preference carryover & \textbf{Placeholder:} session 1 vs session 2 trace diff \\
    Example rationale for auto-approval/check-in & exported traces + screenshots \\
    Verification outcomes & session bundle + trace CSV \\
    \bottomrule
  \end{tabular}
\end{table}

At minimum, the final paper should include one clean exported session bundle and one trace CSV from the demo scenario. Those artifacts support three claims: first, that the system runs end to end on a real task; second, that both model-side and policy-side oversight occur; and third, that traces from one session can affect behavior in the next.

\paragraph{Placeholder for demo video link.} Insert the private video URL as a footnote in the camera-ready or submission draft, as permitted by the CAIS demo CFP.

\paragraph{Placeholder for pilot metrics.} Add 2--4 concrete values from a rehearsed run, for example: number of actions, number of check-ins by initiator, whether the second session reduced unnecessary interruptions, and whether the learned preference appeared in the prompt or preference view.
