\section{Discussion, Limitations, and Next Steps}

Smart Coder is best understood as a governed adaptive-autonomy framework rather than a claim of solved personalization. Its strongest contribution is architectural: it externalizes oversight into a local, inspectable CLI while preserving model-side reasoning as a useful but non-authoritative source of check-ins. This creates a reusable framework for studying how coding-agent autonomy should adapt over time. In that sense, the system complements work on interaction-style modeling and post-action personalization rather than competing directly with it~\\cite{huq2026cowcorpus,liang2026pahf}.

The current system also has clear limitations. The approval policy is heuristic, and its weights are engineering priors rather than learned values. The system does not yet implement interaction-style cold-start priors, reversibility as a distinct risk signal, richer interrupt taxonomy, or asynchronous delegation. Its spec support is also bounded: the model is grounded in a spec digest and asked to report requirement coverage, but Smart Coder does not yet provide a fully structured spec-driven development pipeline.

These limitations are acceptable for a demo paper as long as they are described accurately. The present artifact is a working prototype and study platform, not a finalized personalization method. This is also why the trace infrastructure matters. The artifact is designed to collect the evidence needed to answer the next set of questions: which oversight source is better calibrated, which signals are most predictive, and whether a learned policy can outperform the current heuristic baseline.

The most immediate next step is a controlled lab study using the packaged CLI, exported traces, and a standardized task repository. That study would let us evaluate whether Smart Coder improves the productivity-safety tradeoff relative to static rules or fully manual oversight, and whether users prefer its hybrid milestone-and-risk-driven oversight behavior. Just as importantly, it would let us measure outcomes that recent position work argues are missing from coding-agent evaluation: interruption burden, steerability, verification effort, and calibrated reliance during realistic developer workflows~\\cite{wang2025humansmissing}. It would also let us run shadow-mode comparisons, where alternative policies are evaluated offline on the same traces without changing the live participant experience.

\paragraph{Submission positioning.} For CAIS, we position Smart Coder as both \textbf{(a)} a novel technical contribution in governed adaptive autonomy and \textbf{(b)} a compelling interactive demo of a coding-agent workflow that changes across sessions based on observed developer behavior.

\paragraph{Placeholder for final limitations paragraph.} If the final pilot runs expose concrete rough edges---for example, a miscalibrated check-in or an unnecessary interruption---add one short paragraph here. Reviewers will trust the paper more if the limitations are concrete rather than generic.
