\documentclass[sigconf,review]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}
}

% TODO: Replace with final ACM metadata after acceptance, if required.
\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}

\title{Smart Coder: Dynamic Autonomy for Coding Agents Under Explicit Local Governance}

% TODO: Replace author metadata before submission.
\author{Tanja L. Shukla}
\email{TODO}
\affiliation{%
  \institution{TODO}
  \city{TODO}
  \state{TODO}
  \country{USA}
}

\renewcommand{\shortauthors}{Shukla}

\begin{document}

\begin{abstract}
Coding agents force an uncomfortable tradeoff: either developers approve nearly every action, which slows work, or they grant broad autonomy and accept the risk of subtle bugs, unsafe changes, and scope drift. We present Smart Coder, a governance-first framework for adaptive autonomy in coding agents. Smart Coder treats the language model as untrusted: the model may request reads, declare intent, initiate check-ins, and propose file updates, but it cannot directly read files, write files, or execute commands. A local CLI mediates every side effect using hard constraints, workflow phase, verification outcomes, and interaction traces. The system combines two independent oversight sources: policy-initiated check-ins from the CLI and model-initiated check-ins when the model surfaces uncertainty, assumptions, or architectural tradeoffs. Every interaction is logged and reused to adapt both future policy decisions and future prompt context. The current prototype implements a heuristic baseline with milestone-aware oversight, read/write-split constraints, spec-aware planning, and exportable study traces. We position Smart Coder as both a working research artifact and a framework for studying how adaptive autonomy can better balance developer productivity and safety in agent-assisted software engineering.
\end{abstract}

\keywords{coding agents, adaptive autonomy, human-AI collaboration, software engineering, AI oversight, agentic systems}

\maketitle

\input{sections/01-intro}
\input{sections/02-system}
\input{sections/03-demo}
\input{sections/04-discussion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}
